---
title: "Assignment4"
author: "Kyle Bambling"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Question 1

A common task is to take a set of data that has multiple categorical 
variables and create a table of the number of cases for each combination.
An introductory statistics textbook contains a dataset summarizing student
surveys from several sections of an intro class. The two variables of
interest for us are `Gender` and `Year` which are the students gender and
year in college.

a) Download the dataset and correctly order the `Year` variable using the
        following:

```{r}
Survey <- read.csv('https://www.lock5stat.com/datasets3e/StudentSurvey.csv', na.strings=c('',' ')) %>%
  mutate( Year = factor(Year, levels = c( 'FirstYear', 'Sophomore','Junior','Senior')))
head(Survey)
```

b) Using some combination of `dplyr` functions, produce a data set with
        eight rows that contains the number of responses for each gender:year
        combination. Make sure your table orders the `Year` variable in the
        correct order of `First Year`, `Sophmore`, `Junior`, and then `Senior`.
        
```{r}
survey.2 <- Survey %>%
  count(Year, Sex) %>%
  filter(!is.na(Year))
survey.2
```

c) Using `tidyr` commands, produce a table of the number of responses

```{r}
pivot_wider(survey.2, names_from = Year, values_from = n)
```

## Question 2

From the book website, there is a .csv file of the daily maximum temperature
in Flagstaff at the Pulliam Airport.

```{r}
airport <- read.csv('https://raw.githubusercontent.com/dereksonderegger/444/master/data-raw/FlagMaxTemp.csv')
```

a) Create a line graph that gives the daily maximum temperature for 2005.

```{r}
airport.2005 <- airport %>% filter(Year == 2005)
airport.2005.long <- pivot_longer(airport.2005, X1:X31, names_to = 'Days', values_to = 'Temps')
airport.2005.long <- airport.2005.long %>% 
  mutate( Days = str_replace(Days, pattern='X', replacement='') ) %>%
  drop_na() %>%
  mutate( Date = make_date(year=Year, month=Month, day=Days))

airport.2005.long %>% ggplot( aes(x=Date, y=Temps) ) +
  geom_line()
```

b) Create a line graph that gives the monthly average maximum temperature
        for 2013 - 2015.

```{r}
airport.2013 <- airport %>% filter(2013 <= Year & Year <= 2015)
airport.2013.long <- pivot_longer(airport.2013, X1:X31, names_to = 'Days', values_to = 'Temps')
airport.2013.long <- airport.2013.long %>%
  mutate( Days = str_replace(Days, pattern='X', replacement='') ) %>%
  drop_na() %>%
  group_by(Year, Month) %>%
  summarise( Mean = mean(Temps)) %>%
  mutate( Date = make_date(year=Year, month=Month))

airport.2013.long %>%
  ggplot( aes(x=Date, y=Mean)) +
  geom_line()
  
```
## Question 4

For this problem we will consider two simple data sets.

```{r}
A <- tribble(
  ~Name, ~Car,
  'Alice', 'Ford F150',
  'Bob',   'Tesla Model III',
  'Charlie', 'VW Bug')

B <- tribble(
  ~First.Name, ~Pet,
  'Bob',  'Cat',
  'Charlie', 'Dog',
  'Alice', 'Rabbit')
```

a) Squish the data frames together to generate a data set with three rows
        and three columns. Do two ways: first using `cbind` and then using one
        of the `dplyr` `join` commands.
        
```{r}
B <- rename( B, Name = First.Name )
ABbind <- cbind(A, B)

ABbind <- full_join(A,B)
ABbind
```

b) It turns out that Alice also has a pet guinea pig. Add another row to
        the `B` data set. Do this using either the base function `rbind`, or
        either of the `dplyr` functions `add_row` or `bind_rows`.

```{r}
B <- B %>% add_row( Name = 'Alice', Pet = 'Guinea Pig')
```

c) Squish the `A` and `B` data sets together to generate a data set with
        four rows and three columns. Do this two ways: first using `cbind` and
        then using one of the `dplyr` `join` commands. Which was easier to
        program? Which is more likely to have an error.
        
```{r}
ABjoin <- full_join(A, B)
ABjoin
```